A Survey of Parallelization of Kalman Filters

### 2017-06-10 paper/paper.tex
slowly but surely?

\section{Introduction}
Since their inception in 1960 Kalman filters have been heavily used for the estimation of states in dynamic systems. These filters have numerous applications especially in tracking various objects; they have been used to track object in conjunction with computer vision, radar tracking for airplanes and ships, vehicle and missile guidance systems, orbit calculation for satillites, and particles in HEP. Additionally, they have been used for demographic and economic studies, industrial process control, climate science, and geological surveys. Often these applications, the tracking ones especially, require real-time performance, pushing the performance demands of Kalman filtering.

At first the filtering was done primarily using customized circuits, each designed for the specific application to match the number of states in use. As modern computers developed, new opportunities for flexibility and improved performance were discovered. In many cases parallelism at different levels features in the performance enhancement of the Kalman filter. This survey reviews some of the implementations and applications of parallel Kalman filters.

@@

\section{Background}
Kalman filters \cite{kalman1960new} were originally presented in 1960 to provide a solution to filtering and prediction problems in dynamic systems. In particular, the results targeted signals involved in communication and control systems. Since then, the Kalman filter has been applied to many dynamic systems in fields including physics, signal processing, economics, and others. In this section, we present some background information to enable the reader to better understand the rest of the survey. Interested readers are directed to the following sources: \cite{blackman1986multiple, welch1995introduction, budhiraja2007survey, kalman1960new}.

@@

The Kalman equations are shown in Equations 1 - 5. Equations 1 and 2 are the predict portion, estimating the next state based on past data. Equations 3 - 5 and the correct section, updating the estimated state and Kalman gain based on the measurement and predicted state. There are additional versions such as the Extended Kalman Filter that implement improvements such as time variant error functions, or other changes to improve performance or accuracy.

@@

\section{Modern Implementations and Applications}

@@

\subsection{Cell Tower Power Management}
\cite{rosen2013parallelization}

>>>>

\section{Introduction}
Since their inception in 1960 Kalman filters have been heavily used for the estimation of states in dynamic systems. These filters have numerous applications especially in tracking various objects; they have been used to track objects in conjunction with computer vision, radar tracking for airplanes and ships, vehicle and missile guidance systems, orbit calculation for satellites, and particles in HEP. Additionally, they have been used for demographic and economic studies, industrial process control, climate science, and communication systems. Often these applications, the tracking ones especially, require real-time performance, pushing the performance demands of Kalman filtering.

At first the filtering was done primarily using customized circuits, each designed for the specific application to match the number of states in use. As modern computers developed, new opportunities for flexibility and improved performance were discovered. In many cases parallelism at different levels features in the performance enhancement of the Kalman filter. Commonly the underlying linear algebra calculations are parallelized in one fashion while the broader equations or simultaneous filters are parallelized a different way; however, some of the applications take different approaches as discussed below. In many cases the application area significantly impacts how the filter is parallelized since the number of states varies dramatically between. For this reason we include a discussion of the intended applications alongside the implementations.

@@

\section{Background}
Kalman filters \cite{kalman1960new} were originally presented in 1960 to provide a solution to filtering and prediction problems in dynamic systems. In particular, the results targeted the signal processing involved in communication and control systems. Since then, the Kalman filter has been applied to many dynamic systems in fields including physics, signal processing, economics, and others. In this section, we present some background information to enable the reader to better understand the rest of the survey. Interested readers are directed to the following sources: \cite{blackman1986multiple, welch1995introduction, budhiraja2007survey, kalman1960new}.

@@

The Kalman equations are shown in Equations 1 - 5. Equations 1 and 2 are the predict portion, estimating the next state based on past data. Equations 3 - 5 and the correct section, updating the estimated state and Kalman gain based on the measurement and predicted state. There are additional versions such as the Extended Kalman Filter and Square Root Kalman Filter that implement improvements such as time variant error functions, or other changes to improve performance or accuracy.

@@

\section{Modern Implementations and Applications}
Since the early efforts to parallelize the Kalman filter many advances have been made in computer systems that allow further improvements to the performance. Many of these take a similar approach as Palis and Krecker, using SIMD operations for the matrix calculations and threads or processes for each individual filter. A variety of applications and implementations on modern hardware are present in this section.

@@

\subsection{Cell Phone Management}
Kalman filters play an important role in the processing of MIMO communications such as those used in cell phones. In \cite{rosen2013parallelization}, Rosen et al. present a parallelized implementation of the Kalman filter suitable for modern multicore embedded architectures. The suggested application for this design is for the mapping of cell phone users to cells in the system. This application requires one state per user and all the users over several cells, resulting in hundreds of states.

The authors consider the possibility of separating the equations and running each on a separate core of the processor, but conclude that this idea requires too much communication (particularly of matrix P) to be efficient. Instead, the authors partition the P matrix so that each processor works on all of the equations but only for a section of it. The resulting filter only requires synchronization at the end to bring the results together, so communication among the processors is minimized allowing the authors to maximize communication among cell phones :)

THe authors enable this parallelization by modifying the problem slightly. The full MIMO (multiple input multiple output) problem is broken down into numerous MISO (multiple input single output) problems which limits the number of measurements (possibly to one, the paper wasn't clear) thus eliminating the need to invert a large matrix.
